Downloading the F/RPKM data
---------------------------

Prepare by defining functions etc.
```{r prepare}
library(pheatmap)
library(gplots)
library(ops)
library(calibrate)


do.SVD = function(m, comp.1=1, comp.2=2){ # returns eig.cell
  s <- svd(m)
	ev <- s$d^2 / sum(s$d^2)
	return(s$u[,c(comp.1, comp.2)])
}

project.SVD <- function(m, eig.cell){
	return(t(m) %*% eig.cell)
}

plot.SVD <- function(m, comp.1=1, comp.2=2, groups=rep("blue", ncol(m)), title=""){
	eig <- do.SVD(m, comp.1, comp.2)
	proj <- project.SVD(m, eig)
	xminv <- min(proj[,1]) # - .2 * abs(min(proj[,1]))
	xmaxv <- max(proj[,1]) # + .2 * abs(max(proj[,1]))
	yminv <- min(proj[,2]) # - .2 * abs(min(proj[,2]))
	ymaxv <- max(proj[,2]) # + .2 * abs(max(proj[,2]))
	plot(proj,pch=20,col="white",xlim=c(xminv,xmaxv),ylim=c(yminv,ymaxv),xaxt='n',yaxt='n',xlab="PC1",ylab="PC2",main=title)
	
	points(proj, col=as.character(groups),pch=20) # , #pch=c(rep(15,3),rep(17,3),rep(19,3),rep(18,3),rep(20,2)), cex=2)
	textxy(proj[,1],proj[,2],labs=colnames(m))
}

loadings.SVD <- function(m, comp=1, gene.ids = rownames(m)){
	s <- svd(m)
	l <- s$u[,comp]
	names(l) <- gene.ids
	l.s <- l[order(l)]
	return(l.s)
}

plot.loadings.SVD <- function(m, comp=1, cutoff=0.1, gene.ids = rownames(m)){
	l <- loadings.SVD(m, comp, gene.ids)
	barplot(l[abs(l)>cutoff],las=2,main=paste("PC", comp, "cutoff", cutoff),cex.names=0.6)
}

plotPC <- function(matrix,a,b,desc,colors){
eig <- do.SVD(matrix, a, b)
proj <- project.SVD(matrix, eig)
xminv <- min(proj[,1]) - .2 * abs(min(proj[,1]))
xmaxv <- max(proj[,1]) + .2 * abs(max(proj[,1]))
yminv <- min(proj[,2]) - .2 * abs(min(proj[,2]))
ymaxv <- max(proj[,2]) + .2 * abs(max(proj[,2]))
plot(proj,pch=20,xlim=c(xminv,xmaxv),ylim=c(yminv,ymaxv),xaxt='n',yaxt='n',xlab=paste0("PC",a),ylab=paste("PC",b),col=colors,main=desc)
textxy(proj[,1],proj[,2],labs=rownames(proj))
}

```

Here, we download data from various public sources and extract the brain, heart and kidney samples.

"HPA": Human Protein Atlas
```{r}
#temp <- tempfile()
#download.file(url="http://www.proteinatlas.org/download/rna.csv.zip",destfile=temp)
#hpa <- read.csv(unz(temp, "rna.csv"))
#unlink(temp)

#hpa.heart <- hpa[hpa$Sample=="heart muscle", c("Gene", "Value")]
#hpa.brain <- hpa[hpa$Sample=="cerebral cortex", c("Gene", "Value")]
#hpa.kidney <- hpa[hpa$Sample=="kidney", c("Gene", "Value")]

#hpa.fpkms <- merge(hpa.heart, hpa.brain, by="Gene")
#hpa.fpkms <- merge(hpa.fpkms, hpa.kidney, by="Gene")
#colnames(hpa.fpkms) <- c("ENSG_ID", "HPA_heart", "HPA_brain", "HPA_kidney")
```

Check if the identifiers are unique and write table to file.
```{r}
#length(hpa.fpkms[,1])
#length(unique(hpa.fpkms[,1]))

#write.table(hpa.fpkms,file="hpa_fpkms.txt",quote=F,sep="\t")
```

"Altiso": Alternative isoform regulation in human tissue transcriptomes
```{r}
#temp <- tempfile()
#download.file(url="http://genes.mit.edu/burgelab/Supplementary/wang_sandberg08/hg18.ensGene.CEs.rpkm.txt",destfile=temp)
#altiso <- read.delim(temp, sep="\t")
#unlink(temp)
```

There is no kidney sample here, so just use heart + brain

```{r}
#altiso.fpkms <- altiso[,c("X.Gene","heart","brain")]
#colnames(altiso.fpkms) <- c("ENSG_ID", "AltIso_heart", "AltIso_brain")
```

Check uniqueness of IDs.

```{r}
#length(altiso.fpkms[,1])
#length(unique(altiso.fpkms[,1]))

#write.table(altiso.fpkms,file="altiso_fpkms.txt",quote=F,sep="\t")
```

"GTEx": Genotype-Tissue Expression

This is a big download: 337.8 Mb (as of 2014-02-04)
We also add some code to randomly select one sample from each tissue type; there are many biological replicates in this data set.

```{r}
#temp <- tempfile()
#download.file(url="http://www.broadinstitute.org/gtex/rest/file/download?portalFileId=119363&forDownload=true",destfile=temp)
#header_lines <- readLines(temp, n=2)
#gtex <- read.delim(temp, skip=2, sep="\t")
#unlink(temp)

#write.table(gtex, file="gtex_all.txt", 	quote=F, sep="\t")

#download.file(url="http://www.broadinstitute.org/gtex/rest/file/download?portalFileId=119273&forDownload=true",destfile="GTEx_description.txt")

#metadata <- read.delim("GTEx_description.txt", sep="\t")
```

The metadata table seems to contain entries that are not in the RPKM table.

```{r}
#samp.id <- gsub('-','.',metadata$SAMPID)
#eligible.samples <- which(samp.id %in% colnames(gtex))
#metadata <- metadata[eligible.samples,]
```

Select random heart, kidney and brain samples.

```{r}
#random.heart <- sample(which(metadata$SMTS=="Heart"), size=1)
#random.heart.samplename <- gsub('-','.',metadata[random.heart, "SAMPID"])
#gtex.heart.fpkm <- as.numeric(gtex[,random.heart.samplename])

#random.brain <- sample(which(metadata$SMTS=="Brain"), size=1)
#random.brain.samplename <- gsub('-','.',metadata[random.brain, "SAMPID"])
#gtex.brain.fpkm <- as.numeric(gtex[,random.brain.samplename])

#random.kidney <- sample(which(metadata$SMTS=="Kidney"), size=1)
#random.kidney.samplename <- gsub('-','.',metadata[random.kidney, "SAMPID"])
#gtex.kidney.fpkm <- as.numeric(gtex[,random.kidney.samplename])
```

Get gene IDs on same format as the other data sets by removing the part after the dot; check ID uniqueness and write to file.

```{r}
#gtex.names <- gtex[,"Name"]
#temp_list <- strsplit(as.character(gtex.names), split="\\.")
#gtex.names.nodot <- unlist(temp_list)[2*(1:length(gtex.names))-1]

#gtex.fpkms <- data.frame(ENSG_ID=gtex.names.nodot, GTEx_heart=gtex.heart.fpkm, GTEx_brain=gtex.brain.fpkm,GTEx_kidney=gtex.kidney.fpkm)

#length(gtex.fpkms[,1])
#length(unique(gtex.fpkms[,1]))

#write.table(gtex.fpkms,file="gtex_fpkms.txt",quote=F,sep="\t")
```

*RNA-seq Atlas*

```{r}
#temp <- tempfile()
#download.file(url="http://medicalgenomics.org/rna_seq_atlas/download?download_revision1=1",destfile=temp)
#atlas <- read.delim(temp, sep="\t")
#unlink(temp)

#atlas.fpkms <- atlas[,c("ensembl_gene_id","heart","hypothalamus","kidney")]
#colnames(atlas.fpkms) <- c("ENSG_ID","Atlas_heart","Atlas_brain","Atlas_kidney")
#write.table(atlas.fpkms,file="atlas_fpkms.txt",quote=F,sep="\t")
```

Combining F/RPKM values from public data sets
---------------------------------------------

We will join the data sets on ENSEMBL ID:s, losing a lot of data in the process - but joining on gene symbols or something else would lead to an even worse loss. 

```{r LoadPackagesAndData}
library(org.Hs.eg.db) # for transferring gene identifiers
library(data.table) # for collapsing transcript RPKMs
library(pheatmap) # for nicer visualization
library(edgeR) # for TMM normalization

#hpa.fpkms <- read.delim("hpa_fpkms.txt")
#altiso.fpkms <- read.delim("altiso_fpkms.txt")
#gtex.fpkms <- read.delim("gtex_fpkms.txt")
#atlas.fpkms <- read.delim("atlas_fpkms.txt")
```

The RNA-seq Atlas data set uses many different identifiers, while the other all use ENSG as the primary identifier

Approach 1: Merge on ENSEMBL genes (ENSG) as given in RNA-seq Atlas. Note that there are repeated ENSG ID:s in RNA-seq Atlas, as opposed to the other data sets, so we need to do something about that. In this case, we just sum the transcripts that belong to each ENSG gene. We use data.table for this.

```{r SumTranscripts}
#data.dt <- data.table(atlas.fpkms)
#setkey(data.dt, ENSG_ID)
#temp <- data.dt[, lapply(.SD, sum), by=ENSG_ID]
#collapsed <- as.data.frame(temp)
#atlas.fpkms.summed <- collapsed[,2:ncol(collapsed)] 
#rownames(atlas.fpkms.summed) <- collapsed[,1]

#atlas.fpkms.summed <- atlas.fpkms.summed[2:nrow(atlas.fpkms.summed),]
```

Finally, combine all the data sets into a data frame.

```{r Combine}
#fpkms <- merge(hpa.fpkms, altiso.fpkms, by="ENSG_ID")
#fpkms <- merge(fpkms, gtex.fpkms, by="ENSG_ID")
#fpkms <- merge(fpkms, atlas.fpkms.summed, by.x="ENSG_ID", by.y=0)
#gene_id <- fpkms[,1]
#f <- fpkms[,2:ncol(fpkms)]
#rownames(f) <- gene_id
```

Check how many ENSG IDs we have left.

```{r}
#dim(f)
```

Approach 2: Try to map Entrez symbols to ENSEMBL to recover more ENSG IDs than already present in the table. 

```{r RemapENSEMBL}
#m <- org.Hs.egENSEMBL
#mapped_genes <- mappedkeys(m)
#ensg.for.entrez <- as.list(m[mapped_genes])
#remapped.ensg <- ensg.for.entrez[as.character(atlas$entrez_gene_id)]

#atlas.fpkms$remapped_ensg <- as.character(remapped.ensg)

# And add expression values
#data.dt <- data.table(atlas.fpkms[,2:ncol(atlas.fpkms)])
#setkey(data.dt, remapped_ensg)
#temp <- data.dt[, lapply(.SD, sum), by=remapped_ensg]
#collapsed <- as.data.frame(temp)
#atlas.fpkms.summed <- collapsed[,2:ncol(collapsed)] 
#rownames(atlas.fpkms.summed) <- collapsed[,1]
```

Combine data sets again

```{r Combine2}
#fpkms <- merge(hpa.fpkms, altiso.fpkms, by="ENSG_ID")
#fpkms <- merge(fpkms, gtex.fpkms, by="ENSG_ID")
#fpkms <- merge(fpkms, atlas.fpkms.summed, by.x="ENSG_ID", by.y=0)
#gene_id <- fpkms[,1]
#f <- fpkms[,2:ncol(fpkms)]
#rownames(f) <- gene_id
#write.table(f, file = 'published_rpkms.txt', quote=F)
```

Check how many ENSG IDs we have left.

```{r}
f <- read.delim("published_rpkms.txt",sep=" ")
#dim(f)
```

This looks much better. 
Let's remove all lines where FPKM is close to zero in all samples before we proceed with this version of the data set:

```{r}

f.nozero <- f[-which(rowSums(f[,])<=0.01),]

```

Start by a few correlation heat maps:

**Figure 1B**

Heatmap of Spearman correlations between published expression profiles (# genes = 13,333)

```{r}

pheatmap(cor(f.nozero, method="spearman")) 

```

The brain samples are in a separate cluster, whereas the heart and kidney ones are intermixed.

Alternatively, one could use Pearson correlation (not shown in paper):

```{r}

pheatmap(cor(f.nozero))

```

Sometimes the linear (Pearson) correlation works better on log values.  (not shown in paper):

```{r}
pseudo <- 1
fpkms.log <- log2(f.nozero + pseudo)
pheatmap(cor(fpkms.log))
write.table(fpkms.log, file="published_rpkms_log2.txt", quote=F)
```

What if we drop the genes that have less than FPKM 1 on average? (not shown in paper):

```{r}
f.nolow <- f.nozero[-which(rowMeans(f.nozero)<1),]
pheatmap(cor(log2(f.nolow+pseudo)))
```

Try Anova on a "melted" expression matrix with some metadata:

```{r:anova}
library(reshape)
m <- melt(f.nozero)
colnames(m) <- c("sample_ID","RPKM")
meta <- data.frame(tissue=c("heart","brain","kidney","heart","brain","heart","brain","kidney","heart","brain","kidney"),study=c("HPA","HPA","HPA","AltIso","AltIso","GTex","GTex","GTex","Atlas","Atlas","Atlas"),prep=c(rep("poly-A",8),rep("rRNA-depl",3)),layout=c(rep("PE",3),rep("SE",2),rep("PE",3),rep("SE",3)))
rownames(meta) <- colnames(f.nozero)
tissue <- rep(meta$tissue, each=nrow(f.nozero))
study <- rep(meta$study, each=nrow(f.nozero))
prep <- rep(meta$prep, each=nrow(f.nozero))
layout <- rep(meta$layout, each=nrow(f.nozero))
data <- data.frame(m, tissue=tissue, study=study, prep=prep, layout=layout)

#subset <- data[sample(1:nrow(data), 1000),]
fit <- lm(RPKM ~ prep + layout + study + tissue, data=data)
a <- anova(fit)
maxval = 3000
```

**Figure 1X**

```{r}
barplot(a$"F value"[-5],names.arg=rownames(a)[-5],main="Anova F score, Raw RPKM",ylim=c(0,maxval))
print(a)
```

Let's look at a few SVD plots. 

**Figure 1C**

```{r:pca-fig1d}
colors <- c(1,2,3,1,2,1,2,3,1,2,3)
plotPC(f.nozero, 1, 2, "Published FPKM values \n SVD \n n=13333", colors=colors)
```

The heart samples are clearly separating into their own group.

**Figure 1D** (not included in the current manuscript version)

```{r:pca-fig1e}
plotPC(f.nozero, 2, 3, "Published FPKM values \n SVD \n n=13333", colors=colors)
```

We can plot all pairwise combinations of principal components 1 to 5. (not shown in paper)
Start with SVD on the "raw" F/RPKMs.

```{r}
colors <- c(1,2,3,1,2,1,2,3,1,2,3)

par(mfrow=c(4,4))
for (i in 1:6){
  for(j in 1:6){
		if (i<j){ 
		plotPC(f.nozero,i,j,desc="",colors=colors)
		}
	}
}
```

**Code for figure 2**

Figure 2 deals with log-transformed F/RPKM values from published data sets. 

PCA on log2-FPKM values:

**Figure 2A**

```{r}
pseudo <- 1
fpkms.log <- log2(f.nozero + pseudo)

plotPC(fpkms.log, 1, 2, desc="Published FPKM values, log2 \n SVD \n n=13333", colors=colors)
```

**Figure 2B**

```{r}
plotPC(fpkms.log, 2, 3, desc="Published FPKM values, log2 \n SVD \n n=13333", colors=colors)
```

Alternatively, with PCA (prcomp()) the plot would have looked like this (all combinations of top 5 PCs):

```{r}
colors <- c(1,2,3,1,2,1,2,3,1,2,3)

p <- prcomp(t(fpkms.log))

par(mfrow=c(4,4))
for (i in 1:6){
  for(j in 1:6){
		if (i<j){ 
    	plot(p$x[,i],p$x[,j],pch=20,col=colors,xlab=paste("PC", i),ylab=paste("PC", j))
		}
	}
}

```

**Figure 2C**


Let's have a look at the 100 most highly expressed genes in each sample and see how many of these genes that are shared between the studies


```{r:venn }

library(VennDiagram)

HPA_b <- rownames(f[order(f$HPA_brain,decreasing=T),][1:100,])
HPA_h <- rownames(f[order(f$HPA_heart,decreasing=T),][1:100,])
HPA_k <- rownames(f[order(f$HPA_kidney,decreasing=T),][1:100,])

AltIso_b <- rownames(f[order(f$AltIso_brain,decreasing=T),][1:100,])
AltIso_h <- rownames(f[order(f$AltIso_heart,decreasing=T),][1:100,])

GTEx_b <- rownames(f[order(f$GTEx_brain,decreasing=T),][1:100,])
GTEx_h <- rownames(f[order(f$GTEx_heart,decreasing=T),][1:100,])
GTEx_k <- rownames(f[order(f$GTEx_kidney,decreasing=T),][1:100,])

Atlas_b <- rownames(f[order(f$Atlas_brain,decreasing=T),][1:100,])
Atlas_h <- rownames(f[order(f$Atlas_heart,decreasing=T),][1:100,])
Atlas_k <- rownames(f[order(f$Atlas_kidney,decreasing=T),][1:100,])


```
Let's start with the four brain samples:

```{r:venn brain}

draw.quad.venn(100, 100, 100, 100, 
               length(intersect(HPA_b,AltIso_b)),
               length(intersect(HPA_b,GTEx_b)),
               length(intersect(HPA_b,Atlas_b)),
               length(intersect(AltIso_b,GTEx_b)),
               length(intersect(AltIso_b,Atlas_b)),
               length(intersect(GTEx_b,Atlas_b)),
               length(intersect(intersect(HPA_b,AltIso_b),GTEx_b)),
               length(intersect(intersect(HPA_b,AltIso_b),Atlas_b)),
               length(intersect(intersect(HPA_b,GTEx_b),Atlas_b)),
               length(intersect(intersect(AltIso_b,GTEx_b),Atlas_b)),
               length(intersect(intersect(HPA_b,AltIso_b),intersect(Atlas_b,GTEx_b))),
               category = c("HPA","AltIso","GTEx","Atlas"), lwd = rep(0, 4), lty = rep("solid", 4),
               fill = c("mistyrose","steelblue","lightgoldenrod","darkseagreen")
)

```

Let's investigate the heart samples:

```{r:venn heart}

h_Ids <- list(HPA_h,AltIso_h,GTEx_h,Atlas_h)
draw.quad.venn(100, 100, 100, 100, 
               length(intersect(HPA_h,AltIso_h)),
               length(intersect(HPA_h,GTEx_h)),
               length(intersect(HPA_h,Atlas_h)),
               length(intersect(AltIso_h,GTEx_h)),
               length(intersect(AltIso_h,Atlas_h)),
               length(intersect(GTEx_h,Atlas_h)),
               length(intersect(intersect(HPA_h,AltIso_h),GTEx_h)),
               length(intersect(intersect(HPA_h,AltIso_h),Atlas_h)),
               length(intersect(intersect(HPA_h,GTEx_h),Atlas_h)),
               length(intersect(intersect(AltIso_h,GTEx_h),Atlas_h)),
               length(intersect(intersect(HPA_h,AltIso_h),intersect(Atlas_h,GTEx_h))),
               category = c("HPA","AltIso","GTEx","Atlas"), lwd = rep(0, 4), lty = rep("solid", 4),
               fill = c("mistyrose","steelblue","lightgoldenrod","darkseagreen")
)

```

...and the three kidney samples:

```{r:venn kidney}

k_Ids <- list(HPA_k,GTEx_k,Atlas_k)
draw.triple.venn(100, 100, 100, 
                 length(intersect(HPA_k,GTEx_k)),
                 length(intersect(GTEx_k,Atlas_k)),
                 length(intersect(HPA_k,Atlas_k)),
                 length(intersect(intersect(HPA_k,GTEx_k),Atlas_k)),
                 category = c("HPA","GTEx","Atlas"), lwd = rep(0, 3), lty = rep("solid", 3),
                 fill = c("steelblue","lightgoldenrod","darkseagreen")
)


```

**Figure 2D**

```{r:anova-log}
m <- melt(fpkms.log)
colnames(m) <- c("gene_ID","sample_ID","log2RPKM")
data <- data.frame(m, tissue=tissue, study=study, prep=prep, layout=layout)
#subset <- data[sample(1:nrow(data), 1000),]
fit <- lm(log2RPKM ~ + prep + layout + study + tissue, data=data)
b <- anova(fit)

maxval=3000
barplot(b$"F value"[-5],names.arg=rownames(b)[-5],main="Anova F score, log2-RPKM",ylim=c(0,maxval))
print(b)
```

**Code for figure 3**

Figure 3 is about ComBat.

Combat analysis is performed on logged values (n=13333)


```{r:combat}
library(sva)

meta <- data.frame(study=c(rep("HPA",3),rep("AltIso",2),rep("GTex",3),rep("Atlas",3)),tissue=c("Heart","Brain","Kidney","Heart","Brain","Heart","Brain","Kidney","Heart","Brain","Kidney"))

batch <- meta$study
design <- model.matrix(~as.factor(tissue),data=meta)

combat <- ComBat(dat=fpkms.log,batch=batch,mod=design,numCovs=NULL,par.prior=TRUE)

write.table(combat, file="published_rpkms_combat_log2.txt", quote=F)

pheatmap(cor(combat))

colors <- c(1,2,3,1,2,1,2,3,1,2,3)

plotPC(combat,1,2,colors=colors,desc="Published F/RPKM values, ComBat adjusted log2 TMM values \n SVD \n n=13333")
plotPC(combat,2,3,colors=colors,desc="Published F/RPKM values, ComBat adjusted log2 TMM values \n SVD \n n=13333")

```

Let's have a look again at the 100 most highly expressed genes in each sample and see how many of these genes that are shared between the studies, but this time looking at the values after the ComBat run:

```{r:combatVenn}

HPA_b_c <- rownames(combat[order(combat[,1],decreasing=T),][1:100,])
HPA_h_c <- rownames(combat[order(combat[,2],decreasing=T),][1:100,])
HPA_k_c <- rownames(combat[order(combat[,3],decreasing=T),][1:100,])

AltIso_b_c <- rownames(combat[order(combat[,4],decreasing=T),][1:100,])
AltIso_h_c <- rownames(combat[order(combat[,5],decreasing=T),][1:100,])

GTEx_b_c <- rownames(combat[order(combat[,6],decreasing=T),][1:100,])
GTEx_h_c <- rownames(combat[order(combat[,7],decreasing=T),][1:100,])
GTEx_k_c <- rownames(combat[order(combat[,8],decreasing=T),][1:100,])

Atlas_b_c <- rownames(combat[order(combat[,9],decreasing=T),][1:100,])
Atlas_h_c <- rownames(combat[order(combat[,10],decreasing=T),][1:100,])
Atlas_k_c <- rownames(combat[order(combat[,11],decreasing=T),][1:100,])
```
Brain:
```{r:combatVennBrain}

library(gridExtra)
r <- rectGrob(gp=gpar(fill="white",lwd=0))
top <- arrangeGrob(r, r, r, r, nrow=2)
grid.arrange(top, r, ncol=1, main = "\nBrain Combat", heights=c(2, 1))

draw.quad.venn(100, 100, 100, 100, 
               length(intersect(HPA_b_c,AltIso_b_c)),
               length(intersect(HPA_b_c,GTEx_b_c)),
               length(intersect(HPA_b_c,Atlas_b_c)),
               length(intersect(AltIso_b_c,GTEx_b_c)),
               length(intersect(AltIso_b_c,Atlas_b_c)),
               length(intersect(GTEx_b_c,Atlas_b_c)),
               length(intersect(intersect(HPA_b_c,AltIso_b_c),GTEx_b_c)),
               length(intersect(intersect(HPA_b_c,AltIso_b_c),Atlas_b_c)),
               length(intersect(intersect(HPA_b_c,GTEx_b_c),Atlas_b_c)),
               length(intersect(intersect(AltIso_b_c,GTEx_b_c),Atlas_b_c)),
               length(intersect(intersect(HPA_b_c,AltIso_b_c),intersect(Atlas_b_c,GTEx_b_c))),
               category = c("HPA","AltIso","GTEx","Atlas"), lwd = rep(0, 4), lty = rep("solid", 4),
               fill = c("mistyrose","steelblue","lightgoldenrod","darkseagreen")
)

```
Heart:
```{r:combatVennHeart}

r <- rectGrob(gp=gpar(fill="white",lwd=0))
top <- arrangeGrob(r, r, r, r, nrow=2)
grid.arrange(top, r, ncol=1, main = "\nHeart Combat", heights=c(2, 1))

draw.quad.venn(100, 100, 100, 100, 
               length(intersect(HPA_h_c,AltIso_h_c)),
               length(intersect(HPA_h_c,GTEx_h_c)),
               length(intersect(HPA_h_c,Atlas_h_c)),
               length(intersect(AltIso_h_c,GTEx_h_c)),
               length(intersect(AltIso_h_c,Atlas_h_c)),
               length(intersect(GTEx_h_c,Atlas_h_c)),
               length(intersect(intersect(HPA_h_c,AltIso_h_c),GTEx_h_c)),
               length(intersect(intersect(HPA_h_c,AltIso_h_c),Atlas_h_c)),
               length(intersect(intersect(HPA_h_c,GTEx_h_c),Atlas_h_c)),
               length(intersect(intersect(AltIso_h_c,GTEx_h_c),Atlas_h_c)),
               length(intersect(intersect(HPA_h_c,AltIso_h_c),intersect(Atlas_h_c,GTEx_h_c))),
               category = c("HPA","AltIso","GTEx","Atlas"), lwd = rep(0, 4), lty = rep("solid", 4),
               fill = c("mistyrose","steelblue","lightgoldenrod","darkseagreen")
               
)

```
Kidney:
```{r:combatVennKidney}

r <- rectGrob(gp=gpar(fill="white",lwd=0))
top <- arrangeGrob(r, r, r, r, nrow=2)
grid.arrange(top, r, ncol=1, main = "\n\nKidney Combat", heights=c(2, 1))

draw.triple.venn(100, 100, 100, 
                 length(intersect(HPA_k_c,GTEx_k_c)),
                 length(intersect(GTEx_k_c,Atlas_k_c)),
                 length(intersect(HPA_k_c,Atlas_k_c)),
                 length(intersect(intersect(HPA_k_c,GTEx_k_c),Atlas_k_c)),
                 category = c("HPA","GTEx","Atlas"), lwd = rep(0, 3), lty = rep("solid", 3),
                 fill = c("steelblue","lightgoldenrod","darkseagreen")
)

```

Revisit Anova with combated values.

```{r:anova-combat}

m <- melt(combat)
colnames(m) <- c("sample_ID","combat")
data <- data.frame(m, tissue=tissue, study=study, prep=prep, layout=layout)
fit <- lm(combat ~ + prep + layout + study + tissue, data=data)
c <- anova(fit)

barplot(c$"F value"[-5],names.arg=rownames(b)[-5],main="Anova F score, Combat",ylim=c(0,maxval))

print(c)
```

Let's look at the correlations between the principal components and some sample features

```{r:PC-correlation}

pca <- prcomp(t(f.nozero[,]))

var.percent <- ((pca$sdev)^2)/sum(pca$sdev^2) *100
#var.percent <- (pca$d^2 / sum(pca$d^2) ) *100

barplot(var.percent[1:5], xlab="PC", ylab="Percent Variance",names.arg=1:length(var.percent[1:5]), las=1,ylim=c(0,max(var.percent[1:5])+10), col="gray")


rot <- pca$r
x <- pca$x
plot(pca)
summary(pca)
screeplot(pca,type=c("lines"))


sampleinfo <- read.table("sample_info_published.txt",header=TRUE)

patientsample_NRaw<-cbind(as.character(sampleinfo$Study_labels),
                          as.numeric(sampleinfo$NumberRaw),
                          as.character(sampleinfo$Tissue),
                          as.character(sampleinfo$Preparation),
                          as.character(sampleinfo$readlength),
                          as.numeric(sampleinfo$Numbermapped),
                          as.character(sampleinfo$Readtype))

samplenames <- patientsample_NRaw[,-1]
rownames(samplenames) <- patientsample_NRaw[,1]
pca_comps<-x[,0:7]


Nraw_pca.comp_na <- as.matrix(merge(samplenames,pca_comps,by="row.names",all=TRUE))
Nraw_pca.comp<-na.omit(data.frame(Nraw_pca.comp_na))


#NumberRaw
plot(as.numeric(Nraw_pca.comp[,8]),Nraw_pca.comp[,2],xlab="NRaw",ylab="PCA_C1")

pval_Nraw_pc1<-cor.test(as.numeric(Nraw_pca.comp[,8]),as.numeric(Nraw_pca.comp[,2]),method = "spearman")$p.value
pval_Nraw_pc2<-cor.test(as.numeric(Nraw_pca.comp[,9]),as.numeric(Nraw_pca.comp[,2]),method = "spearman")$p.value
pval_Nraw_pc3<-cor.test(as.numeric(Nraw_pca.comp[,10]),as.numeric(Nraw_pca.comp[,2]),method = "spearman")$p.value
pval_Nraw_pc4<-cor.test(as.numeric(Nraw_pca.comp[,11]),as.numeric(Nraw_pca.comp[,2]),method = "spearman")$p.value
pval_Nraw_pc5<-cor.test(as.numeric(Nraw_pca.comp[,12]),as.numeric(Nraw_pca.comp[,2]),method = "spearman")$p.value
pval_Nraw_pc6<-cor.test(as.numeric(Nraw_pca.comp[,13]),as.numeric(Nraw_pca.comp[,2]),method = "spearman")$p.value
pval_Nraw_pc7<-cor.test(as.numeric(Nraw_pca.comp[,14]),as.numeric(Nraw_pca.comp[,2]),method = "spearman")$p.value

cat(sprintf("Number_of_rawreads~PCAs: PCA1=\"%f\"PCA2=\"%f\"PCA3=\"%f\"PCA4=\"%f\"PCA5=\"%f\"PCA6=\"%f\"\n", pval_Nraw_pc1,pval_Nraw_pc2,pval_Nraw_pc3,pval_Nraw_pc4,pval_Nraw_pc5,pval_Nraw_pc6,pval_Nraw_pc7))
#print(pval_agecorr_pc1)


#NumberMapped
plot(as.numeric(Nraw_pca.comp[,8]),Nraw_pca.comp[,6],xlab="NRaw",ylab="PCA_C1")

pval_Nmap_pc1<-cor.test(as.numeric(Nraw_pca.comp[,8]),as.numeric(Nraw_pca.comp[,6]),method = "spearman")$p.value
pval_Nmap_pc2<-cor.test(as.numeric(Nraw_pca.comp[,9]),as.numeric(Nraw_pca.comp[,6]),method = "spearman")$p.value
pval_Nmap_pc3<-cor.test(as.numeric(Nraw_pca.comp[,10]),as.numeric(Nraw_pca.comp[,6]),method = "spearman")$p.value
pval_Nmap_pc4<-cor.test(as.numeric(Nraw_pca.comp[,11]),as.numeric(Nraw_pca.comp[,6]),method = "spearman")$p.value
pval_Nmap_pc5<-cor.test(as.numeric(Nraw_pca.comp[,12]),as.numeric(Nraw_pca.comp[,6]),method = "spearman")$p.value
pval_Nmap_pc6<-cor.test(as.numeric(Nraw_pca.comp[,13]),as.numeric(Nraw_pca.comp[,6]),method = "spearman")$p.value
pval_Nmap_pc7<-cor.test(as.numeric(Nraw_pca.comp[,14]),as.numeric(Nraw_pca.comp[,6]),method = "spearman")$p.value

cat(sprintf("Number_of_mapped_reads~PCAs: PCA1=\"%f\"PCA2=\"%f\"PCA3=\"%f\"PCA4=\"%f\"PCA5=\"%f\"PCA6=\"%f\"\n", pval_Nmap_pc1,pval_Nmap_pc2,pval_Nmap_pc3,pval_Nmap_pc4,pval_Nmap_pc5,pval_Nmap_pc6,pval_Nmap_pc7))

#Tissue
pval_tissue_pc1<-kruskal.test(as.numeric(Nraw_pca.comp[,8]) ~ as.factor(Nraw_pca.comp[,3]))$p.value
pval_tissue_pc2<-kruskal.test(as.numeric(Nraw_pca.comp[,9]) ~ as.factor(Nraw_pca.comp[,3]))$p.value
pval_tissue_pc3<-kruskal.test(as.numeric(Nraw_pca.comp[,10]) ~ as.factor(Nraw_pca.comp[,3]))$p.value
pval_tissue_pc4<-kruskal.test(as.numeric(Nraw_pca.comp[,11]) ~ as.factor(Nraw_pca.comp[,3]))$p.value
pval_tissue_pc5<-kruskal.test(as.numeric(Nraw_pca.comp[,12]) ~ as.factor(Nraw_pca.comp[,3]))$p.value
pval_tissue_pc6<-kruskal.test(as.numeric(Nraw_pca.comp[,13]) ~ as.factor(Nraw_pca.comp[,3]))$p.value
pval_tissue_pc7<-kruskal.test(as.numeric(Nraw_pca.comp[,14]) ~ as.factor(Nraw_pca.comp[,3]))$p.value

cat(sprintf("Tissues~PCAs: PCA1=\"%f\"PCA2=\"%f\"PCA3=\"%f\"PCA4=\"%f\"PCA5=\"%f\"PCA6=\"%f\"\n", pval_tissue_pc1,pval_tissue_pc2,pval_tissue_pc3,pval_tissue_pc4,pval_tissue_pc5,pval_tissue_pc6,pval_tissue_pc7))

#Library_preparation
pval_tissue_pc1<-kruskal.test(as.numeric(Nraw_pca.comp[,8]) ~ as.factor(Nraw_pca.comp[,4]))$p.value
pval_tissue_pc2<-kruskal.test(as.numeric(Nraw_pca.comp[,9]) ~ as.factor(Nraw_pca.comp[,4]))$p.value
pval_tissue_pc3<-kruskal.test(as.numeric(Nraw_pca.comp[,10]) ~ as.factor(Nraw_pca.comp[,4]))$p.value
pval_tissue_pc4<-kruskal.test(as.numeric(Nraw_pca.comp[,11]) ~ as.factor(Nraw_pca.comp[,4]))$p.value
pval_tissue_pc5<-kruskal.test(as.numeric(Nraw_pca.comp[,12]) ~ as.factor(Nraw_pca.comp[,4]))$p.value
pval_tissue_pc6<-kruskal.test(as.numeric(Nraw_pca.comp[,13]) ~ as.factor(Nraw_pca.comp[,4]))$p.value
pval_tissue_pc7<-kruskal.test(as.numeric(Nraw_pca.comp[,14]) ~ as.factor(Nraw_pca.comp[,4]))$p.value

cat(sprintf("Library_prep~PCAs: PCA1=\"%f\"PCA2=\"%f\"PCA3=\"%f\"PCA4=\"%f\"PCA5=\"%f\"PCA6=\"%f\"\n", pval_tissue_pc1,pval_tissue_pc2,pval_tissue_pc3,pval_tissue_pc4,pval_tissue_pc5,pval_tissue_pc6,pval_tissue_pc7))

#pval_interferoncorr_pc1<-kruskal.test(interferon_pca.comp[complete.cases(interferon_pca.comp),][,2]  ~ interferon_pca.comp[complete.cases(interferon_pca.comp),][,9])$p.value
#sequence_type-paired/single
pval_seqtype_pc1<-kruskal.test(as.numeric(Nraw_pca.comp[,8]) ~ as.factor(Nraw_pca.comp[,7]))$p.value
pval_seqtype_pc2<-kruskal.test(as.numeric(Nraw_pca.comp[,9]) ~ as.factor(Nraw_pca.comp[,7]))$p.value
pval_seqtype_pc3<-kruskal.test(as.numeric(Nraw_pca.comp[,10]) ~ as.factor(Nraw_pca.comp[,7]))$p.value
pval_seqtype_pc4<-kruskal.test(as.numeric(Nraw_pca.comp[,11]) ~ as.factor(Nraw_pca.comp[,7]))$p.value
pval_seqtype_pc5<-kruskal.test(as.numeric(Nraw_pca.comp[,12]) ~ as.factor(Nraw_pca.comp[,7]))$p.value
pval_seqtype_pc6<-kruskal.test(as.numeric(Nraw_pca.comp[,13]) ~ as.factor(Nraw_pca.comp[,7]))$p.value
pval_seqtype_pc7<-kruskal.test(as.numeric(Nraw_pca.comp[,14]) ~ as.factor(Nraw_pca.comp[,7]))$p.value

cat(sprintf("Sequence_type~PCAs: PCA1=\"%f\"PCA2=\"%f\"PCA3=\"%f\"PCA4=\"%f\"PCA5=\"%f\"PCA6=\"%f\"\n", pval_seqtype_pc1,pval_seqtype_pc2,pval_seqtype_pc3,pval_seqtype_pc4,pval_seqtype_pc5,pval_seqtype_pc6,pval_seqtype_pc7))

#readlength- 1x32/1x35/1x76/2x100/2x76 
pval_readlength_pc1<-kruskal.test(as.numeric(Nraw_pca.comp[,8]) ~ as.factor(Nraw_pca.comp[,5]))$p.value
pval_readlength_pc2<-kruskal.test(as.numeric(Nraw_pca.comp[,9]) ~ as.factor(Nraw_pca.comp[,5]))$p.value
pval_readlength_pc3<-kruskal.test(as.numeric(Nraw_pca.comp[,10]) ~ as.factor(Nraw_pca.comp[,5]))$p.value
pval_readlength_pc4<-kruskal.test(as.numeric(Nraw_pca.comp[,11]) ~ as.factor(Nraw_pca.comp[,5]))$p.value
pval_readlength_pc5<-kruskal.test(as.numeric(Nraw_pca.comp[,12]) ~ as.factor(Nraw_pca.comp[,5]))$p.value
pval_readlength_pc6<-kruskal.test(as.numeric(Nraw_pca.comp[,13]) ~ as.factor(Nraw_pca.comp[,5]))$p.value
pval_readlength_pc7<-kruskal.test(as.numeric(Nraw_pca.comp[,14]) ~ as.factor(Nraw_pca.comp[,5]))$p.value

cat(sprintf("readlength~PCAs: PCA1=\"%f\"PCA2=\"%f\"PCA3=\"%f\"PCA4=\"%f\"PCA5=\"%f\"PCA6=\"%f\"\n", pval_readlength_pc1,pval_readlength_pc2,pval_readlength_pc3,pval_readlength_pc4,pval_readlength_pc5,pval_readlength_pc6,pval_readlength_pc7))

