Walkthrough with all data analyses that comes with the manuscript
Comparing published FPKM/RPKM values and reprocessed FPKM values.
========================================================

Prepare by loading packages etc.
```{r prepare}
library(pheatmap)
library(reshape)
library(gplots)
library(ops)
library(calibrate)
library(biomaRt)
library(sva)

```

Data from four different public sources are downloaded and brain, heart and kidney samples are extracted. 

Start with Human Protein Atlas (HPA):

```{r:HPA download}
#temp <- tempfile()
#download.file(url="http://www.proteinatlas.org/download/rna.csv.zip",destfile=temp)
#hpa <- read.csv(unz(temp, "rna.csv"))
#unlink(temp)

#hpa.heart <- hpa[hpa$Sample=="heart muscle", c("Gene", "Value")]
#hpa.brain <- hpa[hpa$Sample=="cerebral cortex", c("Gene", "Value")]
#hpa.kidney <- hpa[hpa$Sample=="kidney", c("Gene", "Value")]

#hpa.fpkms <- merge(hpa.heart, hpa.brain, by="Gene")
#hpa.fpkms <- merge(hpa.fpkms, hpa.kidney, by="Gene")
#colnames(hpa.fpkms) <- c("ENSG_ID", "HPA_heart", "HPA_brain", "HPA_kidney")
```

Check if the identifiers are unique and write table to file.

```{r:HPA check identifiers}
#length(hpa.fpkms[,1])
#length(unique(hpa.fpkms[,1]))

#write.table(hpa.fpkms,file="hpa_fpkms.txt",quote=F,sep="\t")
```

Next dataset is from the article "Alternative isoform regulation in human tissue transcriptomes.." by Wang et.al
AltIso:

```{r:AltIso download}
#temp <- tempfile()
#download.file(url="http://genes.mit.edu/burgelab/Supplementary/wang_sandberg08/hg18.ensGene.CEs.rpkm.txt",destfile=temp)
#altiso <- read.delim(temp, sep="\t")
#unlink(temp)
```

There is no kidney sample here, so just use heart and brain

```{r}
#altiso.fpkms <- altiso[,c("X.Gene","heart","brain")]
#colnames(altiso.fpkms) <- c("ENSG_ID", "AltIso_heart", "AltIso_brain")
```

Check uniqueness of IDs.

```{r:AltIso check identifiers}
#length(altiso.fpkms[,1])
#length(unique(altiso.fpkms[,1]))

#write.table(altiso.fpkms,file="altiso_fpkms.txt",quote=F,sep="\t")
```

Next dataset is derived from "GTEx": Genotype-Tissue Expression

This is a big download: 337.8 Mb (as of 2014-02-04)
We also add some code to randomly select one sample from each tissue type; there are many biological replicates in this data set.

```{r:GTEx download}
#temp <- tempfile()
#download.file(url="http://www.gtexportal.org/home/rest/file/download?portalFileId=175729&forDownload=true",destfile=temp)
#header_lines <- readLines(temp, n=2)
#gtex <- read.delim(temp, skip=2, sep="\t")
#unlink(temp)

#write.table(gtex, file="gtex_all.txt",   quote=F, sep="\t")

#download.file(url="http://www.gtexportal.org/home/rest/file/download?portalFileId=175707&forDownload=true",destfile="GTEx_description.txt")

#metadata <- read.delim("GTEx_description.txt", sep="\t")
```

The metadata table seems to contain entries that are not in the RPKM table.

```{r}
#samp.id <- gsub('-','.',metadata$SAMPID)
#eligible.samples <- which(samp.id %in% colnames(gtex))
#metadata <- metadata[eligible.samples,]
```

Select random heart, kidney and brain samples.

```{r}
#random.heart <- sample(which(metadata$SMTS=="Heart"), size=1)
#random.heart.samplename <- gsub('-','.',metadata[random.heart, "SAMPID"])
#gtex.heart.fpkm <- as.numeric(gtex[,random.heart.samplename])

#random.brain <- sample(which(metadata$SMTS=="Brain"), size=1)
#random.brain.samplename <- gsub('-','.',metadata[random.brain, "SAMPID"])
#gtex.brain.fpkm <- as.numeric(gtex[,random.brain.samplename])

#random.kidney <- sample(which(metadata$SMTS=="Kidney"), size=1)
#random.kidney.samplename <- gsub('-','.',metadata[random.kidney, "SAMPID"])
#gtex.kidney.fpkm <- as.numeric(gtex[,random.kidney.samplename])
```

Get gene IDs on same format as the other data sets by removing the part after the dot; check ID uniqueness and write to file.

```{r}
#gtex.names <- gtex[,"Name"]
#temp_list <- strsplit(as.character(gtex.names), split="\\.")
#gtex.names.nodot <- unlist(temp_list)[2*(1:length(gtex.names))-1]

#gtex.fpkms <- data.frame(ENSG_ID=gtex.names.nodot, GTEx_heart=gtex.heart.fpkm, GTEx_brain=gtex.brain.fpkm,GTEx_kidney=gtex.kidney.fpkm)

#length(gtex.fpkms[,1])
#length(unique(gtex.fpkms[,1]))

#write.table(gtex.fpkms,file="gtex_fpkms.txt",quote=F,sep="\t")
```

*RNA-seq Atlas*

```{r:Atlas download}
#temp <- tempfile()
#download.file(url="http://medicalgenomics.org/rna_seq_atlas/download?download_revision1=1",destfile=temp)
#atlas <- read.delim(temp, sep="\t")
#unlink(temp)

#atlas.fpkms <- atlas[,c("ensembl_gene_id","heart","hypothalamus","kidney")]
#colnames(atlas.fpkms) <- c("ENSG_ID","Atlas_heart","Atlas_brain","Atlas_kidney")
#write.table(atlas.fpkms,file="atlas_fpkms.txt",quote=F,sep="\t")
```

Combining F/RPKM values from public data sets
---------------------------------------------

We will join the data sets on ENSEMBL ID:s, losing a lot of data in the process - but joining on gene symbols or something else would lead to an even worse loss. 

```{r LoadPackagesAndData}
library(org.Hs.eg.db) # for transferring gene identifiers
library(data.table) # for collapsing transcript RPKMs
library(pheatmap) # for nicer visualization
library(edgeR) # for TMM normalization

#hpa.fpkms <- read.delim("hpa_fpkms.txt")
#altiso.fpkms <- read.delim("altiso_fpkms.txt")
#gtex.fpkms <- read.delim("gtex_fpkms.txt")
#atlas.fpkms <- read.delim("atlas_fpkms.txt")
```

The RNA-seq Atlas data set uses many different identifiers, while the other all use ENSG as the primary identifier

Approach 1: Merge on ENSEMBL genes (ENSG) as given in RNA-seq Atlas. Note that there are repeated ENSG ID:s in RNA-seq Atlas, as opposed to the other data sets, so we need to do something about that. In this case, we just sum the transcripts that belong to each ENSG gene. We use data.table for this.

```{r SumTranscripts}
#data.dt <- data.table(atlas.fpkms)
#setkey(data.dt, ENSG_ID)
#temp <- data.dt[, lapply(.SD, sum), by=ENSG_ID]
#collapsed <- as.data.frame(temp)
#atlas.fpkms.summed <- collapsed[,2:ncol(collapsed)] 
#rownames(atlas.fpkms.summed) <- collapsed[,1]

#atlas.fpkms.summed <- atlas.fpkms.summed[2:nrow(atlas.fpkms.summed),]
```

Finally, combine all the data sets into a data frame.

```{r Combine}
#fpkms <- merge(hpa.fpkms, altiso.fpkms, by="ENSG_ID")
#fpkms <- merge(fpkms, gtex.fpkms, by="ENSG_ID")
#fpkms <- merge(fpkms, atlas.fpkms.summed, by.x="ENSG_ID", by.y=0)
#gene_id <- fpkms[,1]
#f <- fpkms[,2:ncol(fpkms)]
#rownames(f) <- gene_id
```

Check how many ENSG IDs we have left.

```{r}
#dim(f)
```

Approach 2: Try to map Entrez symbols to ENSEMBL to recover more ENSG IDs than already present in the table. 

```{r RemapENSEMBL}
#m <- org.Hs.egENSEMBL
#mapped_genes <- mappedkeys(m)
#ensg.for.entrez <- as.list(m[mapped_genes])
#remapped.ensg <- ensg.for.entrez[as.character(atlas$entrez_gene_id)]

#atlas.fpkms$remapped_ensg <- as.character(remapped.ensg)

# And add expression values
#data.dt <- data.table(atlas.fpkms[,2:ncol(atlas.fpkms)])
#setkey(data.dt, remapped_ensg)
#temp <- data.dt[, lapply(.SD, sum), by=remapped_ensg]
#collapsed <- as.data.frame(temp)
#atlas.fpkms.summed <- collapsed[,2:ncol(collapsed)] 
#rownames(atlas.fpkms.summed) <- collapsed[,1]
```

Combine data sets again

```{r Combine2}
#fpkms <- merge(hpa.fpkms, altiso.fpkms, by="ENSG_ID")
#fpkms <- merge(fpkms, gtex.fpkms, by="ENSG_ID")
#fpkms <- merge(fpkms, atlas.fpkms.summed, by.x="ENSG_ID", by.y=0)
#gene_id <- fpkms[,1]
#f <- fpkms[,2:ncol(fpkms)]
#rownames(f) <- gene_id
#write.table(f, file = 'published_rpkms.txt', quote=F)

#instead of downloading everytime:
```
Download the data from local file:

```{r:published download}

published <- read.delim("published_rpkms.txt",sep=" ")

sampleinfo_published <- read.table("sample_info_published.txt",header=TRUE)

```

The published FPKM values are first filtered by removing all lines where FPKM is less or equal to 0.01 in all samples:

```{r:published filter zeros}

published.nozero <- published[-which(rowSums(published[,])<=0.01),]

```

Heatmap of Spearman correlations between published expression profiles (# genes = 13,323), Figure 1b:

```{r:published heatmap spearman}

pheatmap(cor(published.nozero, method="spearman")) 

```

Alternatively, one could use Pearson correlation (not shown in paper):

```{r:published heatmap pearson}

pheatmap(cor(published.nozero))

```

PCA analysis of published FPKM values, Figure 1c:

```{r:published PCA}

colors <- c("indianred", "dodgerblue", "forestgreen",
            "indianred", "dodgerblue",
            "indianred", "dodgerblue", "forestgreen", 
            "indianred", "dodgerblue", "forestgreen")
  
p <- prcomp(t(published.nozero))

shapes <- c(rep(15,3),rep(16,2),rep(17,3),rep(8,3))

plot(p$x[,1],p$x[,2],pch=shapes,cex=1.5,col=colors,xlab=paste("PC1 58% of variance"),ylab=paste("PC2 13% of variance"),main="Published FPKM/RPKM values \n n=13,323")
legend("bottomleft",legend=c("Heart","Brain","Kidney"),col=c("indianred", "dodgerblue", "forestgreen"),cex=1.5,pch=20)
legend("top",legend=c("HPA","AltIso","GTEx","Atlas"),col="black",pch=c(15,16,17,8),ncol=2)
```

We can plot all pairwise combinations of principal components 1 to 5. (not shown in paper):

```{r:published pairwise PCA}

par(mfrow=c(4,4))
for (i in 1:6){
  for(j in 1:6){
  	if (i<j){ 
		plot(p$x[,i],p$x[,j],pch=20,col=colors,xlab=paste("PC",i),ylab=paste("PC",j),main="Published FPKM values \n n=13,323")
		}
	}
}

```

Look a bit closer at PCs 1-3 in prcomp (not shown in paper):

```{r:published PC 1-3}

      #PC1:
      load.pc1 <- p$rotation[,1][order(p$rotation[,1])]
      extreme.pc1 <- c(tail(load.pc1), head(load.pc1))

      extreme.pc1.ensg <- names(extreme.pc1)
      ensembl = useMart("ensembl", dataset = "hsapiens_gene_ensembl") #select the ensembl database
      extreme.pc1.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc1.ensg,
                           mart=ensembl)
      
      q <- extreme.pc1.symbols[,2]
      names(q) <- extreme.pc1.symbols[,1]

      fpkm.pc1 <- cbind(q[extreme.pc1.ensg],published.nozero[extreme.pc1.ensg,])
      names(fpkm.pc1)[names(fpkm.pc1) == 'q[extreme.pc1.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc1, names.arg=q[extreme.pc1.ensg],las=2,main="Genes w highest absolute loadings in PC1 (raw RPKM)")
      
      print(fpkm.pc1)

      #PC2:
      load.pc2 <- p$rotation[,2][order(p$rotation[,2])]
      extreme.pc2 <- c(tail(load.pc2), head(load.pc2))
      
      extreme.pc2.ensg <- names(extreme.pc2)
      extreme.pc2.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc2.ensg,
                           mart=ensembl)
      
      q <- extreme.pc2.symbols[,2]
      names(q) <- extreme.pc2.symbols[,1]

      fpkm.pc2 <- cbind(q[extreme.pc2.ensg],published.nozero[extreme.pc2.ensg,])
      names(fpkm.pc2)[names(fpkm.pc2) == 'q[extreme.pc2.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc2, names.arg=q[extreme.pc2.ensg],las=2,main="Genes w highest absolute loadings in PC2 (raw RPKM)")
      
      print(fpkm.pc2)

      

      #PC3:
      load.pc3 <- p$rotation[,3][order(p$rotation[,3])]
      extreme.pc3 <- c(tail(load.pc3), head(load.pc3))
      
      extreme.pc3.ensg <- names(extreme.pc3)
      extreme.pc3.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc3.ensg,
                           mart=ensembl)
      
      q <- extreme.pc3.symbols[,2]
      names(q) <- extreme.pc3.symbols[,1]


      fpkm.pc3 <- cbind(q[extreme.pc3.ensg],published.nozero[extreme.pc3.ensg,])
      names(fpkm.pc3)[names(fpkm.pc3) == 'q[extreme.pc3.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc3, names.arg=q[extreme.pc3.ensg],las=2,main="Genes w highest absolute loadings in PC3 (raw RPKM)")
      
      print(fpkm.pc3)

```
             
Try Anova on a "melted" expression matrix with some metadata, Figure 1d:

```{r:published anova}

m <- melt(published.nozero)
colnames(m) <- c("sample_ID","RPKM")

meta <- sampleinfo_published[,c("Study","Tissue","Preparation","NumberRaw","Numbermapped","Readtype")]
rownames(meta) <- colnames(published.nozero)
tissue <- rep(meta$Tissue, each=nrow(published.nozero))
study <- rep(meta$Study, each=nrow(published.nozero))
prep <- rep(meta$Preparation, each=nrow(published.nozero))
layout <- rep(meta$Readtype, each=nrow(published.nozero))
raw <- rep(meta$NumberRaw, each=nrow(published.nozero))
mapped <- rep(meta$Numbermapped, each=nrow(published.nozero))
data <- data.frame(m, tissue=tissue, study=study, prep=prep, layout=layout,nraw=raw,nmapped=mapped)
fit <- lm(RPKM ~ layout + prep + nraw + study + tissue, data=data)
a <- anova(fit)
maxval = 100

barplot(100*a$"Sum Sq"[1:5]/sum(a$"Sum Sq"[1:5]),names.arg=rownames(a[1:5,]),main="Anova, published FPKM/RPKM values",ylim=c(0,maxval))

```

Try log2 transformation of the published FPKM values:

```{r:published log tranform}

pseudo <- 1
published.log <- log2(published.nozero + pseudo)

```

Heatmap of Spearman correlations between published expression profiles with log2 values, Figure 2a:

```{r:published log heatmap spearman}

pheatmap(cor(published.log),method="spearman")

```

PCA analysis of log2 published FPKM values, Figure 2b:

```{r:published log PCA 1&2}

colors <- c("indianred", "dodgerblue", "forestgreen",
            "indianred", "dodgerblue",
            "indianred", "dodgerblue", "forestgreen", 
            "indianred", "dodgerblue", "forestgreen")

p.log <- prcomp(t(published.log))

shapes <- c(rep(15,3),rep(16,2),rep(17,3),rep(8,3))

plot(p.log$x[,1],p.log$x[,2],pch=shapes,col=colors,xlab=paste("PC1 31% of variance"),ylab=paste("PC2 27% of variance"),main="log2 Published FPKM/RPKM values \n n=13,323")
legend("bottomright",legend=c("Heart","Brain","Kidney"),col=c("indianred", "dodgerblue", "forestgreen"),cex=1.5,pch=20,bty="n")
legend("top",legend=c("HPA","AltIso","GTEx","Atlas"),col="black",pch=c(15,16,17,8),ncol=2)


```

Figure 2c:

```{r:published log PCA 2&3}

plot(p.log$x[,2],p.log$x[,3],pch=shapes,col=colors,xlab=paste("PC2 27% of variance"),ylab=paste("PC3 19% of variance"),main="log2 Published FPKM values \n n=13,323")
legend("bottomright",legend=c("Heart","Brain","Kidney"),col=c("indianred", "dodgerblue", "forestgreen"),cex=1.5,pch=20,bty="n")
legend("topright",legend=c("HPA","AltIso","GTEx","Atlas"),col="black",pch=c(15,16,17,8),ncol=2)


```

Cross-validattion by leaving out one of the studies, in this case AltIso, Figure 2d: 

```{r:leave-one-out-pca}

p.loo <- published.log[,-c(4,5)]
colors.loo <- colors[-c(4,5)]
p.loo.log <- prcomp(t(p.loo))

p.add <- published.log[,c(4,5)]
projection <- t(p.add) %*% p.loo.log$rotation

p.original.plus.new <- rbind(p.loo.log$x, projection)
col.original.plus.new <- c(colors.loo, colors[c(4,5)])
plot(p.original.plus.new[,2],p.original.plus.new[,3],pch=c(rep(15,3),rep(17,3),rep(8,3),rep(22,nrow(projection))),col=col.original.plus.new,xlab="PC2",ylab="PC3",main="log2 Published FPKM/RPKM values; AltIso projected onto existing PCs \n n=13,323",xlim=c(-150,100))

legend("bottomleft",legend=c("Heart","Brain","Kidney"),col=c("indianred", "dodgerblue", "forestgreen"),cex=1.5,pch=20,bty="n")
legend("top",legend=c("HPA","GTEx","Atlas","AltIso"),col="black",pch=c(15,17,8,22),ncol=2)


```

We can plot all pairwise combinations of principal components 1 to 5. (not shown in paper):

```{r:published log PCA pairwise}

par(mfrow=c(4,4))
for (i in 1:6){
  for(j in 1:6){
    if (i<j){ 
		plot(p.log$x[,i],p.log$x[,j],pch=shapes,col=colors,xlab=paste("PC",i),ylab=paste("PC",j),main="log2 Published FPKM values \n n=13323")
		}
	}
}

```

Look a bit closer at PCs 1-3 in prcomp (not shown in paper):

```{r:published log PC 1-3}
     
     #PC1:
     load.pc1 <- p.log$rotation[,1][order(p.log$rotation[,1])]
     extreme.pc1 <- c(tail(load.pc1), head(load.pc1))

     extreme.pc1.ensg <- names(extreme.pc1)
     extreme.pc1.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc1.ensg,
                           mart=ensembl)
     
      q <- extreme.pc1.symbols[,2]
      names(q) <- extreme.pc1.symbols[,1]

      fpkm.log.pc1 <- cbind(q[extreme.pc1.ensg],published.nozero[extreme.pc1.ensg,])
      names(fpkm.log.pc1)[names(fpkm.log.pc1) == 'q[extreme.pc1.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc1, names.arg=q[extreme.pc1.ensg],las=2,main="Genes w highest absolute loadings in PC1 (log2 RPKM)")
      
      print(fpkm.log.pc1)

      #PC2:
      load.pc2 <- p.log$rotation[,2][order(p.log$rotation[,2])]
      extreme.pc2 <- c(tail(load.pc2), head(load.pc2))
      
      extreme.pc2.ensg <- names(extreme.pc2)
      extreme.pc2.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc2.ensg,
                           mart=ensembl)

      q <- extreme.pc2.symbols[,2]
      names(q) <- extreme.pc2.symbols[,1]
      
      fpkm.log.pc2 <- cbind(q[extreme.pc2.ensg],published.nozero[extreme.pc2.ensg,])
      names(fpkm.log.pc2)[names(fpkm.log.pc2) == 'q[extreme.pc2.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc2, names.arg=q[extreme.pc2.ensg],las=2,main="Genes w highest absolute loadings in PC2 (log2 RPKM)")
      
      print(fpkm.log.pc2)
      
      #PC3:
      load.pc3 <- p.log$rotation[,3][order(p$rotation[,3])]
      extreme.pc3 <- c(tail(load.pc3), head(load.pc3))
      
      extreme.pc3.ensg <- names(extreme.pc3)
      extreme.pc3.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc3.ensg,
                           mart=ensembl)
      
      q <- extreme.pc3.symbols[,2]
      names(q) <- extreme.pc3.symbols[,1]
      
      fpkm.log.pc3 <- cbind(q[extreme.pc3.ensg],published.nozero[extreme.pc3.ensg,])
      names(fpkm.log.pc3)[names(fpkm.log.pc3) == 'q[extreme.pc3.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc3, names.arg=q[extreme.pc3.ensg],las=2,main="Genes w highest absolute loadings in PC3 (log2 RPKM)")
      
      print(fpkm.log.pc3)


```

To further validate the above results, indicating that tissue specificity appears mainly in PC 2 and 3, we will extract the 500 genes with highest loadings in each component and plot the corresponding published FPKM values in a heatmap (not shown in paper):

```{r:published log top 500 loadings heatmap}
    
     load.pc1 <- abs(p.log$rotation[,1])[order(abs(p.log$rotation[,1]),decreasing=TRUE)]
     top.pc1 <- names(load.pc1[1:500]) 

     load.pc2 <- abs(p.log$rotation[,2])[order(abs(p.log$rotation[,2]),decreasing=TRUE)]
     top.pc2 <- names(load.pc2[1:500])

     load.pc3 <- abs(p.log$rotation[,3])[order(abs(p.log$rotation[,3]),decreasing=TRUE)]
     top.pc3 <- names(load.pc3[1:500])

     pheatmap(cor(published[top.pc1,]),method="spearman")
     pheatmap(cor(published[top.pc2,]),method="spearman")
     pheatmap(cor(published[top.pc3,]),method="spearman")
```

Try Anova on a "melted" expression matrix with logged values and some metadata, Figure 2e:

```{r:published log anova}
n <- melt(published.log)
colnames(n) <- c("sample_ID","RPKM")
meta <- sampleinfo_published[,c("Study","Tissue","Preparation","NumberRaw","Numbermapped","Readtype")]
rownames(meta) <- colnames(published.log)
tissue <- rep(meta$Tissue, each=nrow(published.log))
study <- rep(meta$Study, each=nrow(published.log))
prep <- rep(meta$Preparation, each=nrow(published.log))
layout <- rep(meta$Readtype, each=nrow(published.log))
raw <- rep(meta$NumberRaw, each=nrow(published.log))
mapped <- rep(meta$Numbermapped, each=nrow(published.log))
data <- data.frame(n, tissue=tissue, study=study, prep=prep, layout=layout,nraw=raw,nmapped=mapped)
fit <- lm(RPKM ~ layout + prep + nraw + study + tissue, data=data)
b <- anova(fit)
maxval = 100

barplot(100*b$"Sum Sq"[1:5]/sum(b$"Sum Sq"[1:5]),names.arg=rownames(a[1:5,]),main="Anova, log2 published FPKM/RPKM values",ylim=c(0,maxval))
```

Combat analysis is performed on log2 values (n=13,323):

```{r:published log combat}

meta <- data.frame(study=c(rep("HPA",3),rep("AltIso",2),rep("GTex",3),rep("Atlas",3)),tissue=c("Heart","Brain","Kidney","Heart","Brain","Heart","Brain","Kidney","Heart","Brain","Kidney"))

batch <- meta$study
design <- model.matrix(~1,data=meta)
combat <- ComBat(dat=published.log,batch=batch,mod=design,numCovs=NULL,par.prior=TRUE)

```

Heatmap of Spearman correlations between published expression profiles after combat run (# genes = 13,323), Figure 3a:

```{r:published log combat heatmap spearman}

pheatmap(cor(combat, method="spearman")) 

```

PCA analysis of published FPKM values after combat run, Figure 3b:

```{r:published log combat PCA}

colors <- c("indianred", "dodgerblue", "forestgreen",
            "indianred", "dodgerblue",
            "indianred", "dodgerblue", "forestgreen", 
            "indianred", "dodgerblue", "forestgreen")

p.combat <- prcomp(t(combat))

shapes <- c(rep(15,3),rep(16,2),rep(17,3),rep(8,3))

plot(p.combat$x[,1],p.combat$x[,2],pch=shapes,col=colors,xlab=paste("PC1 54% of variance"),ylab=paste("PC2 38% of variance"),main="Published FPKM values \n COMBAT \n n=13,323")
legend("bottomright",legend=c("Heart","Brain","Kidney"),col=c("indianred", "dodgerblue", "forestgreen"),cex=1.5,pch=20,bty="n")
legend("topright",legend=c("HPA","AltIso","GTEx","Atlas"),col="black",pch=c(15,16,17,8),ncol=2)



```

We can plot all pairwise combinations of principal components 1 to 5. (not shown in paper):

```{r:published log combat PCA pairwise}

par(mfrow=c(4,4))
for (i in 1:6){
  for(j in 1:6){
    if (i<j){ 
		plot(p.combat$x[,i],p.combat$x[,j],pch=shapes,col=colors,xlab=paste("PC",i),ylab=paste("PC",j),main="Published FPKM values \n COMBAT \ n=13323")
		}
	}
}

```

Look a bit closer at PCs 1-3 in prcomp (not shown in paper):

```{r:published log combat PC 1-3}
    
     #PC1:
     load.pc1 <- p.combat$rotation[,1][order(p.combat$rotation[,1])]
     extreme.pc1 <- c(tail(load.pc1), head(load.pc1))

     extreme.pc1.ensg <- names(extreme.pc1)
     extreme.pc1.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc1.ensg,
                           mart=ensembl)

      q <- extreme.pc1.symbols[,2]
      names(q) <- extreme.pc1.symbols[,1]

      fpkm.combat.pc1 <- cbind(q[extreme.pc1.ensg],published.nozero[extreme.pc1.ensg,])
      names(fpkm.combat.pc1)[names(fpkm.combat.pc1) == 'q[extreme.pc1.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc1, names.arg=q[extreme.pc1.ensg],las=2,main="Genes w highest absolute loadings in PC1 (ComBat log2 RPKM)")
      
      print(fpkm.combat.pc1)
      
      #PC2:
      load.pc2 <- p.combat$rotation[,2][order(p.combat$rotation[,2])]
      extreme.pc2 <- c(tail(load.pc2), head(load.pc2))

      extreme.pc2.ensg <- names(extreme.pc2)
      extreme.pc2.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc2.ensg,
                           mart=ensembl)
      q <- extreme.pc2.symbols[,2]
      names(q) <- extreme.pc2.symbols[,1]

      fpkm.combat.pc2 <- cbind(q[extreme.pc2.ensg],published.nozero[extreme.pc2.ensg,])
      names(fpkm.combat.pc2)[names(fpkm.combat.pc2) == 'q[extreme.pc2.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc2, names.arg=q[extreme.pc2.ensg],las=2,main="Genes w highest absolute loadings in PC2 (ComBat log2 RPKM)")
      
      print(fpkm.combat.pc2)
      
      #PC3:
      load.pc3 <- p.combat$rotation[,3][order(p.combat$rotation[,3])]
      extreme.pc3 <- c(tail(load.pc3), head(load.pc3))

      extreme.pc3.ensg <- names(extreme.pc3)
      extreme.pc3.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc3.ensg,
                           mart=ensembl)

      q <- extreme.pc3.symbols[,2]
      names(q) <- extreme.pc3.symbols[,1]

      fpkm.combat.pc3 <- cbind(q[extreme.pc3.ensg],published.nozero[extreme.pc3.ensg,])
      names(fpkm.combat.pc3)[names(fpkm.combat.pc3) == 'q[extreme.pc3.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc3, names.arg=q[extreme.pc3.ensg],las=2,main="Genes w highest absolute loadings in PC3 (ComBat log2 RPKM)")
      
      print(fpkm.combat.pc3)
    
```

Revisit Anova with combated values, Figure 3c:

```{r:published log combat anova}

o <- melt(combat)
colnames(o) <- c("sample_ID","combat")
meta <- sampleinfo_published[,c("Study","Tissue","Preparation","NumberRaw","Numbermapped","Readtype")]
rownames(meta) <- colnames(combat)
tissue <- rep(meta$Tissue, each=nrow(combat))
study <- rep(meta$Study, each=nrow(combat))
prep <- rep(meta$Preparation, each=nrow(combat))
layout <- rep(meta$Readtype, each=nrow(combat))
raw <- rep(meta$NumberRaw, each=nrow(combat))
mapped <- rep(meta$Numbermapped, each=nrow(combat))
data <- data.frame(o, tissue=tissue, study=study, prep=prep, layout=layout,nraw=raw,nmapped=mapped)
fit <- lm(combat ~ layout + prep + nraw + study + tissue, data=data)
c <- anova(fit)
maxval = 100

barplot(100*c$"Sum Sq"[1:5]/sum(c$"Sum Sq"[1:5]),names.arg=rownames(c[1:5,]),main="Anova Combat",ylim=c(0,maxval))

```

Comparing FPKMs for FASTQ files reprocessed with TopHat and Cufflinks:

```{r:cufflinks download data}

cufflinks <- read.delim("fpkm_table_tophat.txt")

sampleinfo_cufflinks <- read.delim("sample_info_reprocessed.txt")

```

First, we will restrict the data set to only include protein coding genes using the ensembl based R package biomaRt:

```{r:cufflinks filter protein coding}

gene_ids <- as.vector(cufflinks[,1])

ensembl = useMart("ensembl", dataset = "hsapiens_gene_ensembl") #select the ensembl database

gene_type <- getBM(attributes=c("ensembl_gene_id", "gene_biotype"), 
                   filters = "ensembl_gene_id",
                   values=gene_ids,
                   mart=ensembl)

pc <- subset(gene_type[,1],gene_type[,2]=="protein_coding")

cufflinks_pc <- cufflinks[match(pc,cufflinks[,1]),]

```

Let's remove all lines where FPKM is close to zero in all samples before we proceed with this version of the data set:

```{r:cufflinks filter zeros}

cufflinks_pc_nozero <- cufflinks_pc[-which(rowSums(cufflinks_pc[,3:16])<=0.01),]

```

Heatmap of Spearman correlations between reprocessed expression profiles (# genes = 19,475), Figure 4a:

```{r:cufflinks heatmap spearman}

pheatmap(cor(cufflinks_pc_nozero[,3:16], method="spearman")) 

```

Let's look at a few PCA plots, Figure 4b:

```{r:cufflinks PCA}

cufflinks_fpkms <- cufflinks_pc_nozero[,3:16]
rownames(cufflinks_fpkms) <- cufflinks_pc_nozero[,1]

p.cufflinks <- prcomp(t(cufflinks_fpkms))

colors <- c("dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred")          

shapes_cufflinks <- c(rep(11,3),rep(8,3),rep(17,3),rep(15,3),rep(16,2))

plot(p.cufflinks$x[,1],p.cufflinks$x[,2],pch=shapes_cufflinks,col=colors,xlab=paste("PC1 87% of variance"),ylab=paste("PC2 7.7% of variance"),main="Reprocessed FPKM values \n n=19,475")
legend("bottomright",legend=c("Heart","Brain","Kidney"),col=c("indianred", "dodgerblue", "forestgreen"),cex=1.5,pch=20,bty="n")
legend("top",legend=c("EoGE","Atlas","BodyMap","HPA","AltIso"),col="black",pch=c(11,8,17,15,16),ncol=2)



```

We can plot all pairwise combinations of principal components 1 to 5 (not shown in paper):

```{r:cufflinks PCA pairwise}

colors <- c("dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred")

par(mfrow=c(4,4))
for (i in 1:6){
  for(j in 1:6){
    if (i<j){ 
      plot(p.cufflinks$x[,i],p.cufflinks$x[,j],pch=shapes_cufflinks,col=colors,xlab=paste("PC",i),ylab=paste("PC",j),main="Cufflinks FPKM values \n n=19475")
		}
	}
}

```

Look at PCA loadings for PC1-3 (not shown in paper):

```{r:cufflinks PC 1-3}
      
      #PC1:
      load.pc1 <- p.cufflinks$rotation[,1][order(p.cufflinks$rotation[,1])]
      extreme.pc1 <- c(tail(load.pc1), head(load.pc1))
      
      extreme.pc1.ensg <- names(extreme.pc1)
      extreme.pc1.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc1.ensg,
                           mart=ensembl)
      
      q <- extreme.pc1.symbols[,2]
      names(q) <- extreme.pc1.symbols[,1]
      
      fpkm.cuff.pc1 <- cbind(q[extreme.pc1.ensg],cufflinks_fpkms[extreme.pc1.ensg,])
      names(fpkm.cuff.pc1)[names(fpkm.cuff.pc1) == 'q[extreme.pc1.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc1, names.arg=q[extreme.pc1.ensg],las=2,main="Genes w highest absolute loadings in PC1 (raw Cufflinks FPKM)")
      
      print(fpkm.cuff.pc1)
      
      #PC2:
      load.pc2 <- p.cufflinks$rotation[,2][order(p.cufflinks$rotation[,2])]
      extreme.pc2 <- c(tail(load.pc2), head(load.pc2))
      
      extreme.pc2.ensg <- names(extreme.pc2)
      extreme.pc2.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc2.ensg,
                           mart=ensembl)
      
      q <- extreme.pc2.symbols[,2]
      names(q) <- extreme.pc2.symbols[,1]
      
      fpkm.cuff.pc2 <- cbind(q[extreme.pc2.ensg],cufflinks_fpkms[extreme.pc2.ensg,])
      names(fpkm.cuff.pc2)[names(fpkm.cuff.pc2) == 'q[extreme.pc2.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc2, names.arg=q[extreme.pc2.ensg],las=2,main="Genes w highest absolute loadings in PC2 (raw Cufflinks FPKM)")
      
      print(fpkm.cuff.pc2)
      
      #PC3:
      load.pc3 <- p.cufflinks$rotation[,3][order(p.cufflinks$rotation[,3])]
      extreme.pc3 <- c(tail(load.pc3), head(load.pc3))
      
      extreme.pc3.ensg <- names(extreme.pc3)
      extreme.pc3.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc3.ensg,
                           mart=ensembl)
      
      q <- extreme.pc3.symbols[,2]
      names(q) <- extreme.pc3.symbols[,1]
      
      fpkm.cuff.pc3 <- cbind(q[extreme.pc3.ensg],cufflinks_fpkms[extreme.pc3.ensg,])
      names(fpkm.cuff.pc3)[names(fpkm.cuff.pc3) == 'q[extreme.pc3.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc3, names.arg=q[extreme.pc3.ensg],las=2,main="Genes w highest absolute loadings in PC3 (raw Cufflinks FPKM)")
      
      print(fpkm.cuff.pc3)
      
```
Mitochondrially encoded genes have relatively high expresseion levels, FPKM values of several thousands.

Anova analysis of different batch factors, Figure 4c:

```{r:cufflinks anova}

p <- melt(cufflinks_fpkms)
colnames(p) <- c("sample_ID","Cuff_FPKM")
meta <- sampleinfo_cufflinks[,c("Study","Tissue","Preparation","NumberRaw","Numbermapped","Readtype")]
rownames(meta) <- colnames(cufflinks_fpkms)
tissue <- rep(meta$Tissue, each=nrow(cufflinks_fpkms))
study <- rep(meta$Study, each=nrow(cufflinks_fpkms))
prep <- rep(meta$Preparation, each=nrow(cufflinks_fpkms))
layout <- rep(meta$Readtype, each=nrow(cufflinks_fpkms))
nraw <- rep(meta$NumberRaw, each=nrow(cufflinks_fpkms))
data <- data.frame(p, tissue=tissue, study=study, prep=prep, layout=layout, nraw=nraw)
fit <- lm(Cuff_FPKM ~ layout + prep + nraw + study + tissue, data=data)
d <- anova(fit)
maxval = 100

barplot(100*d$"Sum Sq"[1:5]/sum(d$"Sum Sq"[1:5]),names.arg=rownames(d[1:5,]),main="Anova, Cufflinks FPKM",ylim=c(0,maxval))

```

Try log2 transformation of the reprocessed FPKM values:

```{r:cufflinks log transform}

pseudo <- 1
cufflinks_log <- log2(cufflinks_fpkms + pseudo)

```

Heatmap of Spearman correlations between log2 reprocessed cufflinks FPKM values, Figure 4d:

```{r:cufflinks log heatmap spearman }

pheatmap(cor(cufflinks_log) ,method="spearman")

```

PCA analysis of log2 reprocessed cufflinks FPKM values, Figure 4e:

```{r:cufflinks log PCA 1&2}

colors <- c("dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred")          

p.log.cufflinks <- prcomp(t(cufflinks_log))

shapes_cufflinks <- c(rep(11,3),rep(8,3),rep(17,3),rep(15,3),rep(16,2))
plot(p.log.cufflinks$x[,1],p.log.cufflinks$x[,2],pch=shapes_cufflinks,col=colors,xlab=paste("PC1 33% of variance"),ylab=paste("PC2 25% of variance"),main="log2 reprocessed cufflinks FPKM values \n n=19,475")
legend("bottomleft",legend=c("Heart","Brain","Kidney"),col=c("indianred", "dodgerblue", "forestgreen"),cex=1.5,pch=20,bty="n")
legend("top",legend=c("EoGE","Atlas","BodyMap","HPA","AltIso"),col="black",pch=c(11,8,17,15,16),ncol=2)


```

Figure 4f:

```{r:cufflinks log PCA 2&3}

plot(p.log.cufflinks$x[,2],p.log.cufflinks$x[,3],pch=shapes_cufflinks,col=colors,xlab=paste("PC2 25% of variance"),ylab=paste("PC3 22% of variance"),main="log2 reprocessed cufflinks FPKM values \n n=19,475")
legend("bottomleft",legend=c("Heart","Brain","Kidney"),col=c("indianred", "dodgerblue", "forestgreen"),cex=1.5,pch=20,bty="n")
legend("top",legend=c("EoGE","Atlas","BodyMap","HPA","AltIso"),col="black",pch=c(11,8,17,15,16),ncol=2)


```

We can plot all pairwise combinations of principal components 1 to 5. (not shown in paper):

```{r:cufflinks log PCA pairwise}

par(mfrow=c(4,4))
for (i in 1:6){
  for(j in 1:6){
    if (i<j){ 
  	plot(p.log.cufflinks$x[,i],p.log.cufflinks$x[,j],pch=shapes_cufflinks,col=colors,xlab=paste("PC",i),ylab=paste("PC",j),main="log2 reprocessed FPKM values \n n=19475")
		}
	}
}

```

Cross-validattion by leaving out one of the studies, in this case AltIso, Figure 4g: 

```{r:leave-one-out-pca cufflinks}

colors <- c("dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred", "forestgreen",
            "dodgerblue", "indianred") 

p.loo <- cufflinks_log[,-c(13,14)]
colors.loo <- colors[-c(13,14)]
p.loo.log <- prcomp(t(p.loo))

p.add <- cufflinks_log[,c(13,14)]
projection <- t(p.add) %*% p.loo.log$rotation

p.original.plus.new <- rbind(p.loo.log$x, projection)
col.original.plus.new <- c(colors.loo, colors[c(13,14)])
plot(p.original.plus.new[,2],p.original.plus.new[,3],pch=c(shapes_cufflinks[1:12],rep(22,nrow(projection))),col=col.original.plus.new,xlab="PC2",ylab="PC3",main="log2 Cufflinks FPKM values; AltIso projected onto existing PCs \n n=19,475,",xlim=c(-150,100))

legend("bottomleft",legend=c("Heart","Brain","Kidney"),col=c("indianred", "dodgerblue", "forestgreen"),cex=1.5,pch=20,bty="n")
legend("topleft",legend=c("HPA","GTEx","Atlas","AltIso"),col="black",pch=c(11,8,17,15,22),ncol=2)


```

Look a bit closer at PCs 1-3 in prcomp for the logged FPKM values from cufflinks (not shown in paper):

```{r:cufflinks log PC 1-3}
      
      #PC1
      load.pc1 <- p.log.cufflinks$rotation[,1][order(p.log.cufflinks$rotation[,1])]
      extreme.pc1 <- c(tail(load.pc1), head(load.pc1))
      
      extreme.pc1.ensg <- names(extreme.pc1)
      extreme.pc1.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc1.ensg,
                           mart=ensembl)
      
      q <- extreme.pc1.symbols[,2]
      names(q) <- extreme.pc1.symbols[,1]
      
      fpkm.logcuff.pc1 <- cbind(q[extreme.pc1.ensg],cufflinks_fpkms[extreme.pc1.ensg,])
      names(fpkm.logcuff.pc1)[names(fpkm.logcuff.pc1) == 'q[extreme.pc1.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc1, names.arg=q[extreme.pc1.ensg],las=2,main="Genes w highest absolute loadings in PC1 (log2 Cufflinks FPKM)")
      
      print(fpkm.combat.pc1)
            
      #PC2
      load.pc2 <- p.log.cufflinks$rotation[,2][order(p.log.cufflinks$rotation[,2])]
      extreme.pc2 <- c(tail(load.pc2), head(load.pc2))
      
      extreme.pc2.ensg <- names(extreme.pc2)
      extreme.pc2.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc2.ensg,
                           mart=ensembl)
      
      q <- extreme.pc2.symbols[,2]
      names(q) <- extreme.pc2.symbols[,1]
      
      fpkm.logcuff.pc2 <- cbind(q[extreme.pc2.ensg],cufflinks_fpkms[extreme.pc2.ensg,])
      names(fpkm.logcuff.pc2)[names(fpkm.logcuff.pc2) == 'q[extreme.pc2.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc2, names.arg=q[extreme.pc2.ensg],las=2,main="Genes w highest absolute loadings in PC2 (log2 Cufflinks FPKM)")
      
      print(fpkm.combat.pc2)
      
      #PC3
      load.pc3 <- p.log.cufflinks$rotation[,3][order(p.log.cufflinks$rotation[,3])]
      extreme.pc3 <- c(tail(load.pc3), head(load.pc3))
      
      extreme.pc3.ensg <- names(extreme.pc3)
      extreme.pc3.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc3.ensg,
                           mart=ensembl)

      q <- extreme.pc3.symbols[,2]
      names(q) <- extreme.pc3.symbols[,1]
  
      fpkm.logcuff.pc3 <- cbind(q[extreme.pc3.ensg],cufflinks_fpkms[extreme.pc3.ensg,])
      names(fpkm.logcuff.pc3)[names(fpkm.logcuff.pc3) == 'q[extreme.pc3.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc3, names.arg=q[extreme.pc3.ensg],las=2,main="Genes w highest absolute loadings in PC3 (log2 Cufflinks FPKM)")
      
      print(fpkm.combat.pc3)
    
```

Seems to yield a heart vs. brain separation

To further validate the above results, indicating that tissue specificity appears mainly in PC 3, we will extract the 500 genes with highest loadings in each component and plot the corresponding cufflinks FPKM values in a heatmap (not shown in paper):

```{r:cufflinks log top 500 loadings heatmap}
     
     cufflinks_values <- cufflinks_pc_nozero[,3:16]
     rownames(cufflinks_values) <- cufflinks_pc_nozero[,1]

     load.pc1 <- abs(p.log.cufflinks$rotation[,1])[order(abs(p.log.cufflinks$rotation[,1]),decreasing=TRUE)]
     top.pc1 <- names(load.pc1[1:500])
     load.pc2 <- abs(p.log.cufflinks$rotation[,2])[order(abs(p.log.cufflinks$rotation[,2]),decreasing=TRUE)]
     top.pc2 <- names(load.pc2[1:500])
     load.pc3 <- abs(p.log.cufflinks$rotation[,3])[order(abs(p.log.cufflinks$rotation[,3]),decreasing=TRUE)]
     top.pc3 <- names(load.pc3[1:500])
     
     pheatmap(cor(cufflinks_values[top.pc1,]),method="spearman")
     pheatmap(cor(cufflinks_values[top.pc2,]),method="spearman")
     pheatmap(cor(cufflinks_values[top.pc3,]),method="spearman")

```
    
Try Anova on a "melted" expression matrix with logged cufflinks values and some metadata (not shown in paper):

```{r:cufflinks log anova}
q <- melt(cufflinks_log[,])
colnames(q) <- c("sample_ID","logFPKM")
meta <- sampleinfo_cufflinks[,c("Study","Tissue","Preparation","NumberRaw","Numbermapped","Readtype")]
rownames(meta) <- colnames(cufflinks_log)
tissue <- rep(meta$Tissue, each=nrow(cufflinks_log))
study <- rep(meta$Study, each=nrow(cufflinks_log))
prep <- rep(meta$Preparation, each=nrow(cufflinks_log))
layout <- rep(meta$Readtype, each=nrow(cufflinks_log))
data <- data.frame(q, tissue=tissue, study=study, prep=prep, layout=layout)
fit <- lm(logFPKM ~ layout + prep + nraw + study + tissue, data=data)
e <- anova(fit)
maxval = 100

barplot(100*e$"Sum Sq"[1:5]/sum(e$"Sum Sq"[1:5]),names.arg=rownames(e[1:5,]),main="Anova, Cufflinks log2 FPKM",ylim=c(0,maxval))
```

Combat analysis for removal of batch effects (n=19,475):

```{r:cufflinks log combat}

meta <- data.frame(study=c(rep("EoGE",3),rep("Atlas",3),rep("BodyMap",3),rep("HPA",3),rep("AltIso",2)),tissue=c("Brain","Heart","Kidney","Brain","Heart","Kidney","Brain","Heart","Kidney","Brain","Heart","Kidney","Brain","Heart"),prep=c(rep("poly-A",3),rep("rRNA-depl",3),rep("poly-A",8)),layout=c(rep("PE",3),rep("SE",3),rep("PE",6),rep("SE",2)))

batch <- meta$study
design <- model.matrix(~1,data=meta)
combat.cufflinks <- ComBat(dat=cufflinks_log,batch=batch,mod=design,numCovs=NULL,par.prior=TRUE)
rownames(combat.cufflinks) <- rownames(cufflinks_log)

```

Heatmap of Spearman correlations between reprocessed cufflinks FPKM values after ComBat run, Figure 4h:

```{r:cufflinks log combat heatmap spearman}

pheatmap(cor(combat.cufflinks),method="spearman")

```

PCA analysis on reprocessed cufflinks FPKM values after ComBat run, Figure 4i:

```{r:cufflinks log combat PCA}

p.combat.cufflinks <- prcomp(t(combat.cufflinks))

shapes_cufflinks <- c(rep(11,3),rep(8,3),rep(17,3),rep(15,3),rep(16,2))


plot(p.combat.cufflinks$x[,1],p.combat.cufflinks$x[,2],pch=shapes_cufflinks,col=colors,xlab=paste("PC1 54% of variance"),ylab=paste("PC2 37% of variance"),main="Cufflinks FPKM values \n COMBAT \n n=19,475")
legend("bottomleft",legend=c("Heart","Brain","Kidney"),col=c("indianred", "dodgerblue", "forestgreen"),cex=1.5,pch=20,bty="n")
legend("top",legend=c("EoGE","Atlas","BodyMap","HPA","AltIso"),col="black",pch=c(11,8,17,15,16),ncol=2)

```

We can plot all pairwise combinations of principal components 1 to 5. (not shown in paper):

```{r:cufflinks log combat PCA pairwise}

par(mfrow=c(4,4))
for (i in 1:6){
  for(j in 1:6){
    if (i<j){ 
  	plot(p.combat.cufflinks$x[,i],p.combat.cufflinks$x[,j],pch=shapes_cufflinks,col=colors,xlab=paste("PC",i),ylab=paste("PC",j),main="Cufflinks FPKM values \n COMBAT \ n=19,475")
		}
	}
}

```

Look a bit closer at PCs 1-3 in prcomp (not shown in paper):

```{r:cufflinks log combat PC 1-3}
      #PC1:
      load.pc1 <- p.combat.cufflinks$rotation[,1][order(p.combat.cufflinks$rotation[,1])]
      extreme.pc1 <- c(tail(load.pc1), head(load.pc1))
      
      extreme.pc1.ensg <- names(extreme.pc1)
      extreme.pc1.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc1.ensg,
                           mart=ensembl)
      
      q <- extreme.pc1.symbols[,2]
      names(q) <- extreme.pc1.symbols[,1]

      fpkm.combatcuff.pc1 <- cbind(q[extreme.pc1.ensg],cufflinks_fpkms[extreme.pc1.ensg,])
      names(fpkm.combatcuff.pc1)[names(fpkm.combatcuff.pc1) == 'q[extreme.pc1.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc1, names.arg=q[extreme.pc1.ensg],las=2,main="Genes w highest absolute loadings in PC1 (COMBAT Cufflinks FPKM)")
      
      print(fpkm.combatcuff.pc1)


      #PC2:
      load.pc2 <- p.combat.cufflinks$rotation[,2][order(p.combat.cufflinks$rotation[,2])]
      extreme.pc2 <- c(tail(load.pc2), head(load.pc2))
      
      extreme.pc2.ensg <- names(extreme.pc2)
      extreme.pc2.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc2.ensg,
                           mart=ensembl)
      
      q <- extreme.pc2.symbols[,2]
      names(q) <- extreme.pc2.symbols[,1]
      
      fpkm.combatcuff.pc2 <- cbind(q[extreme.pc2.ensg],cufflinks_fpkms[extreme.pc2.ensg,])
      names(fpkm.combatcuff.pc2)[names(fpkm.combatcuff.pc2) == 'q[extreme.pc2.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc2, names.arg=q[extreme.pc2.ensg],las=2,main="Genes w highest absolute loadings in PC2 (COMBAT Cufflinks FPKM)")
      
      print(fpkm.combatcuff.pc2)

      #PC3:      
      load.pc3 <- p.combat.cufflinks$rotation[,3][order(p.combat.cufflinks$rotation[,3])]
      extreme.pc3 <- c(tail(load.pc3), head(load.pc3))
      
      extreme.pc3.ensg <- names(extreme.pc3)
      extreme.pc3.symbols <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol"), 
                           filters = "ensembl_gene_id",
                           values=extreme.pc3.ensg,
                           mart=ensembl)
      
      q <- extreme.pc3.symbols[,2]
      names(q) <- extreme.pc3.symbols[,1]

      fpkm.combatcuff.pc3 <- cbind(q[extreme.pc3.ensg],cufflinks_fpkms[extreme.pc3.ensg,])
      names(fpkm.combatcuff.pc3)[names(fpkm.combatcuff.pc3) == 'q[extreme.pc3.ensg]'] <- 'Gene Symbol'
      
      barplot(extreme.pc3, names.arg=q[extreme.pc3.ensg],las=2,main="Genes w highest absolute loadings in PC3 (COMBAT Cufflinks FPKM)")
      
      print(fpkm.combatcuff.pc3)

      
      
```

Revisit ANOVA analysis on reprocessed Cufflinks FPKM values after ComBat run, Figure 4j:

```{r:cufflinks log combat anova}

q <- melt(combat.cufflinks)
colnames(q) <- c("sample_ID","combat")
meta <- sampleinfo_cufflinks[,c("Study","Tissue","Preparation","NumberRaw","Numbermapped","Readtype")]
rownames(meta) <- colnames(combat.cufflinks)
tissue <- rep(meta$Tissue, each=nrow(combat.cufflinks))
study <- rep(meta$Study, each=nrow(combat.cufflinks))
prep <- rep(meta$Preparation, each=nrow(combat.cufflinks))
layout <- rep(meta$Readtype, each=nrow(combat.cufflinks))
raw <- rep(meta$NumberRaw, each=nrow(combat.cufflinks))
mapped <- rep(meta$Numbermapped, each=nrow(combat.cufflinks))
data <- data.frame(q, tissue=tissue, study=study, prep=prep, layout=layout,nraw=raw,nmapped=mapped)
fit <- lm(combat ~ prep + nraw + layout + study + tissue, data=data)
f <- anova(fit)
maxval = 100
barplot(100*f$"Sum Sq"[1:5]/sum(f$"Sum Sq"[1:5]),names.arg=rownames(f[1:5,]),main="Anova Cufflinks Combat",ylim=c(0,maxval))


```